Lightweight Real-Time Feature Monitoring

### 2020-07-17 paper.tex
....

% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf]{acmart}
\usepackage{listings}
\usepackage{dirtytalk}
\usepackage[shortlabels]{enumitem}
\usepackage{amssymb}
\usepackage{minitoc}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{mathtools}

@@

\section{Method}
\label{sec:Method}
A data stream is a continuous collection of events, each timestamped and with multiple fields. In this paper, we refer to the event's fields as features. Our goal is to detect data pattern deviations for each feature in a streaming fashion. To that end, we devise a method that works with two windows and comprises two phases, a batch analysis phase and a stream analysis one. The batch analysis is performed on a static in time window which we call the reference window. The stream analysis is performed over an emulated sliding window that groups data stream events. We aggregate the contents of the reference window with an exact histogram and the sliding window with an approximate histogram aggregation based on Exponential Moving Averages (EMAs). We then apply a statistical distribution divergence function to both histogram aggregations to obtain a divergence value. In the batch phase, we build the distribution of these values so we can define a percentile threshold for the newly measured divergence values in streaming. We raise alerts for features above certain percentiles, \textit{e.g.}, the 99th percentile.

@@

            \ForEach {$count \in counts$} \Comment{//EMA suppression}
                \State $count \gets count \times 2^{- \frac{1}{n_{1/2}}}$
            \EndFor
        \EndFunction
    \end{algorithmic}
\end{algorithm}
This procedure is illustrated in Figure \ref{fig:EMA-hist-build}.
\begin{figure}[!htb]
    \begin{center}
    %\hspace{6cm}
      \includegraphics[scale=0.6]{figures/ema-hist.png}
      \caption[]{Reference equal-height histogram}
      \label{fig:EMA-hist-build}
    \end{center}
\end{figure}

The EMA-based approximate histogram emulates a sliding window as an exponential tail. This aggregation shows constant memory and time complexity regarding window size, making it ideal for our lightweight real-time system.

@@

\subsubsection{Building the Distribution of Divergence Values} \label{sec:sampling-batch}
Secondly, we aim to build the distribution of expected distance or divergence values, to later on threshold the observed distance values during the streaming phase. To that end, we make $S$ samples of transactions, each with the same tuple-based size of the sliding window we emulate in streaming. Each sample is a contiguous block of transactions, thus preserving the order of transactions and the time-dependency property of a time-series. With this sampling we aim to build a distribution of divergence values and reliably estimate the upper tail of that distribution. In \cite{SAMM}, the authors show how many samples must be made to have at least one value in the upper tail of the distribution. Defining $\alpha$ as our threshold and $\frac{\alpha}{n_{features}}$ as our threshold corrected by the Holm Bonferroni multiple test correction, $\gamma$ is defined as the probability of \textbf{not} having at least one point in the $\frac{\alpha}{n_{features}}$ upper tail of the distribution and the minimum number of samples $S$ to make is given by:
\begin{equation*}
    \label{eq:optimal-n-samples}
    \textit{S} = \ceil[\bigg]{\frac{\log\gamma}{log(1 - \frac{\alpha}{n_{features}})}}
\end{equation*}
For each sample, and each feature, we compute the approximated EMA histogram using the bins from the reference histogram. We also add two extra bins that will cover the region to the left and the right of our histogram, respectively. This way, observed values outside of the reference period will be placed in these special bins, either left or right. We compute multiple sample histograms to encode distributions over smaller time periods that may exist within the large reference period. We use an approximated histogram to mimic the target approximated histogram we will incrementally maintain in the streaming environment.

For each feature, we have now \textit{S} histograms, one for each sample. We compute the divergence between each of the \textit{S} histograms and the reference histogram. Figure \ref{fig:compute-sample-distances} illustrates this process as we apply our divergence or distance function for each sample's EMA-based histogram and the reference histogram, obtaining \textit{S} divergence measurements.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.4]{figures/compute-sample-distances.png}
      \caption[Compute sample's distance values]{Compute divergence values between reference and samples}
      \label{fig:compute-sample-distances}
    \end{center}
\end{figure}
These $S$ measurements are divergence measurements between random samples of data and the reference period. Hence, we claim we now know the distribution of expected distance values. Given a new distance value measured between another random window and the reference window, we can compute its probability and produce alerts if the probability is below a certain threshold. Note that the histograms for each sample are EMA-based histograms just like the emulated sliding window histogram. This ensures a certain degree of fidelity in the test we make in streaming because we measure the divergence between a random sample of data (our streaming histogram or any of the sample's histogram) and the reference one.

\subsubsection{Burn-In Period at System Initialization}
When we boot up our system in the streaming phase we initially have empty histogram aggregations. To avoid this burn-in period, we propose saving the last sample's histogram for each feature so that the sliding window EMA-based histogram aggregation is initialized with it.

\subsubsection{Batch Phase Artifacts} \label{sec:batch-artifacts-summary}
By the end of the batch phase, for each feature, we have: 
\begin{enumerate}
    \item a reference histogram
    \item a distribution of divergence values
    \item the last sample's histogram
\end{enumerate}

@@

We periodically perform a multi-feature change detection test. The test is done by first computing the divergence between the reference histogram for a feature \textit{x} and the sliding window EMA-based histogram of the same feature. This value is tested against the known distribution of distance values for feature \textit{x}. We test to see if the percentile of the computed distance relative to the distribution of divergence is above a certain percentile. The probability of a divergence value $val_x$ (or \textit{p-value}) is:
\[pvalue(val_x) = 1 - percentile(val_x)\]
For example, if we define alerts to have 1\% or less probability to happen, it means the system will raise an alert for distance values above the 99th-percentile or \textit{p-value} = 0.01. We alert a feature \textit{x} as divergent at a given timestamp if for a given probability threshold $\alpha$:
\[ percentile(val_x) > 1 - \alpha \]
or:
\begin{equation}
    \label{eq:alert-test}
    pvalue(val_x) < \alpha
\end{equation}

\subsubsection{Multiple Test Correction} \label{sec:multi-test}
Note that we periodically apply this test (Equation \ref{eq:alert-test}) for each feature, effectively repeating the analysis. When considering several hypotheses, the problem of multiplicity arises \cite{MultiTestProblem-Pubmed, MultiTestProblem-Dickhaus2014, MultiTestProblem-Miller1966, multiple-test-correction-geoffrey}: the more hypotheses are checked, the higher the probability of obtaining false positives. In the literature, many multiple test correction methods are proposed, but we opted to use the Holm-Bonferroni multiple test correction because it is a simple and easy to compute correction that can reduce the number of false negatives while having as target to control the number of false positives. Furthermore, it does not require to assume any independence between the tested hypothesis (as other methods do) which would almost surely be a wrong assumption for realistic datasets. In statistics, the Family-Wise Error Rate (FWER) \cite{MultivariateMT, MultitestTamhane2018AdvancesIP} is the probability of having one or more false positives when performing multiple hypotheses tests, \textit{i.e.}, considering one hypothesis true when it is not. The Holm-Bonferroni method \cite{HolmBonferroni} is one of many approaches for controlling the Family-Wise Error Rate \textit{(FWER)} by adjusting the rejection criteria for each hypothesis.

\subsubsection*{Holm-Bonferroni Correction} \label{sec:holmbonferroni}
The Holm-Bonferroni correction takes Equation \ref{eq:alert-test} and corrects the right-side threshold $\alpha$ for each feature. This correction is done by dividing the threshold $\alpha$ by a correction factor. First, we compute the \textit{p-value} of each feature. We then order our features by \textit{p-value}, in ascending order. Next, starting from the first feature, the one with the smallest \textit{p-value}, and proceeding until the end of the list of features, we check if:
\begin{equation}
    pvalue(val_x) < \alpha / (m - k)
    \label{eq:corrected-pvalue}
\end{equation}
with $\alpha$ as the probability to reject the null hypothesis, \textit{m} as the number of hypotheses to test (in our case the number of features) and \textit{k} as the 0-based index of the current feature. We stop at the first feature that fails this test, and every feature processed up until this one is considered to be a divergent feature and alerts are raised.

\section{Experiments}
\label{sec:Experiments}

Much scientist

\section{Discussion}
\label{sec:Discussion}

Fight

>>>>

% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf]{acmart}
\usepackage{listings}
\usepackage{dirtytalk}
\usepackage[shortlabels]{enumitem}
\usepackage{amssymb}
\usepackage{minitoc}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{mathtools}
\usepackage{subcaption}

@@

\section{Method}
\label{sec:Method}
A data stream is a continuous collection of events, each timestamped and with multiple fields. In this paper, we refer to the event's fields as features. Our goal is to detect data pattern deviations for each feature in a streaming fashion. To that end, we devise a method that works with two windows and comprises two phases, a batch analysis phase and a stream analysis one. The batch analysis is performed on a static in time window which we call the reference window. The stream analysis is performed over a exponential tail which emulates a sliding window that groups data stream events. We aggregate the contents of the reference window with an exact histogram and the sliding window with an approximate histogram aggregation based on Exponential Moving Averages (EMAs). We then apply a statistical distribution divergence function to both histogram aggregations to obtain a divergence value. In the batch phase, we build the distribution of these values so we can define a percentile threshold for the newly measured divergence values in streaming. We raise alerts for features above certain percentiles, \textit{e.g.}, the 99th percentile.

@@

            \ForEach {$count \in counts$} \Comment{//EMA suppression}
                \State $count \gets count \times 2^{- \frac{1}{n_{1/2}}}$
            \EndFor
        \EndFunction
    \end{algorithmic}
\end{algorithm}
This procedure is illustrated in Figure \ref{fig:EMA-hist-build}.
\begin{figure}[!htb]
    \begin{center}
    %\hspace{6cm}
      \includegraphics[scale=0.6]{figures/ema-hist.png}
      \caption[]{Reference equal-height histogram}
      \label{fig:EMA-hist-build}
    \end{center}
\end{figure}
The EMA-based approximate histogram emulates a sliding window as an exponential tail, seen in Figure \ref{fig:EMA-tail}. This aggregation shows constant memory and time complexity regarding window size, making it ideal for our lightweight real-time system.
\begin{figure}[!htb]
    \begin{center}
    %\hspace{6cm}
      \includegraphics[scale=0.35]{figures/emavswindow.pdf}
      \caption[]{EMA tail versus sliding window weights}
      \label{fig:EMA-tail}
    \end{center}
\end{figure}

@@

\subsubsection{Building the Distribution of Divergence Values} \label{sec:sampling-batch}
Secondly, we aim to build the distribution of expected distance or divergence values, to later on threshold the observed distance values during the streaming phase. To that end, we make $S$ samples of transactions, each with the same tuple-based size of the sliding window we emulate in streaming. Each sample is a contiguous block of transactions, thus preserving the order of transactions and the time-dependency property of a time-series. With this sampling we aim to build a distribution of divergence values and reliably estimate the upper tail of that distribution. In \cite{SAMM}, the authors show how many samples must be made to have at least one value in the upper tail of the distribution. Defining $\alpha$ as our threshold (and $\frac{\alpha}{n_{features}}$ as our threshold corrected by the Holm-Bonferroni multiple test correction \cite{HolmBonferroni}, detailed in Section \ref{sec:multi-test}), $\gamma$ is defined as the probability of \textbf{not} having at least one point in the $\frac{\alpha}{n_{features}}$ upper tail of the distribution and the minimum number of samples $S$ to make is given by:
\begin{equation*}
    \label{eq:optimal-n-samples}
    \textit{S} = \ceil[\bigg]{\frac{\log\gamma}{log(1 - \frac{\alpha}{n_{features}})}}
\end{equation*}

For each sample, and each feature, we compute the EMA histogram using the bins from the reference histogram. We also add two extra bins that will cover the region to the left and the right of our histogram, respectively. This way, observed values outside of the reference period will be placed in these special bins, either left or right. We compute multiple sample histograms to encode distributions over smaller time periods that may exist within the large reference period. We use an EMA histogram to mimic the target EMA histogram we will incrementally maintain in the streaming environment.

For each feature, we have now \textit{S} histograms, one for each sample. We compute the divergence between each of the \textit{S} histograms and the reference histogram. Figure \ref{fig:compute-sample-distances} illustrates this process as we apply our divergence or distance function for each sample's EMA-based histogram and the reference histogram, obtaining \textit{S} divergence measurements.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.4]{figures/compute-sample-distances.png}
      \caption[Compute sample's distance values]{Compute divergence values between reference and samples}
      \label{fig:compute-sample-distances}
    \end{center}
\end{figure}
These $S$ measurements are divergence measurements between random samples of data and the reference period. Hence, we claim we now know the distribution of expected distance values. Given a new distance value measured between another random window and the reference window, we can compute its probability and produce alerts if the probability is below a certain threshold. Note that the histograms for each sample are EMA-based histograms just like the streaming histogram. This ensures a certain degree of fidelity in the test we make in streaming because we measure the divergence between a random sample of data (our streaming histogram or any of the sample's histogram) and the reference one.

\subsubsection{Burn-In Period at System Initialization}
When we boot up our system in the streaming phase we initially have empty histogram aggregations. To avoid this burn-in period, we propose saving the last sample's histogram for each feature so that the streaming EMA-based histogram aggregation is initialized with it.

\subsubsection{Batch Phase Artifacts} \label{sec:batch-artifacts-summary}
By the end of the batch phase, for each feature, we have: 
\begin{enumerate}
    \item a reference histogram
    \item the distribution of divergence values
    \item the last sample's histogram
\end{enumerate}

@@

We periodically perform a multi-feature change detection test. The test is done by first computing the divergence (here denoted as $div_x$) between the reference histogram for a feature \textit{x} and the current streaming EMA histogram for $x$. This value is tested against the known distribution of distance values for feature \textit{x}. We test to see if the percentile of the computed distance relative to the distribution of divergence is above a certain percentile. The probability of a divergence value $div_x$ is:
\[pvalue(div_x) = 1 - percentile(div_x)\]
For example, if we define alerts to have 1\% or less probability to happen, it means the system will raise an alert for distance values above the 99th-percentile or below a probability of \textit{p-value} = 0.01. We alert a feature \textit{x} as divergent at a given timestamp if for a given probability threshold $\alpha$:
\[ percentile(div_x) > 1 - \alpha \]
or:
\begin{equation}
    \label{eq:alert-test}
    pvalue(div_x) < \alpha
\end{equation}

\subsubsection{Multiple Test Correction} \label{sec:multi-test}
Note that we periodically apply this test (Equation \ref{eq:alert-test}) for each feature, effectively repeating the analysis. When considering several hypotheses, the problem of multiplicity arises \cite{MultiTestProblem-Pubmed, MultiTestProblem-Dickhaus2014, MultiTestProblem-Miller1966, multiple-test-correction-geoffrey}: the more hypotheses are checked, the higher the probability of obtaining false positives. In the literature, many multiple test correction methods are proposed, but we opted to use the Holm-Bonferroni \cite{HolmBonferroni} multiple test correction because it is a simple and easy to compute correction that can reduce the number of false negatives while having as target to control the number of false positives. Furthermore, it does not require to assume any independence between the tested hypothesis (as other methods do) which would almost surely be a wrong assumption for realistic datasets. In statistics, the Family-Wise Error Rate (FWER) \cite{MultivariateMT, MultitestTamhane2018AdvancesIP} is the probability of having one or more false positives when performing multiple hypotheses tests, \textit{i.e.}, considering one hypothesis true when it is not. The Holm-Bonferroni method is one of many approaches for controlling the Family-Wise Error Rate \textit{(FWER)} by adjusting the rejection criteria for each hypothesis. The Holm-Bonferroni correction takes Equation \ref{eq:alert-test} and corrects the right-side threshold $\alpha$ for each feature. This correction is done by dividing the threshold $\alpha$ by a correction factor. First, we compute the \textit{p-value} of each feature. We then order our features by \textit{p-value}, in ascending order. Next, starting from the first feature, the one with the smallest \textit{p-value}, and proceeding until the end of the list of features, we check if:
\begin{equation}
    pvalue(val_x) < \alpha / (m - k)
    \label{eq:corrected-pvalue}
\end{equation}
with $\alpha$ as the probability to reject the null hypothesis, \textit{m} as the number of hypotheses to test (in our case the number of features) and \textit{k} as the 0-based index of the current feature. We stop at the first feature that fails this test, and every feature processed up until this one is considered to be a divergent feature and an alert per feature is logged with corresponding probability value.

\section{Experiments with Synthetic Data}
\label{sec:SyntheticExperiments}
In order to test our proposed system, we generated multiple synthetic datasets before moving on to real data.

\subsection{Single-Feature Analysis}
As an initial step, we started by validating our system using single-feature datasets before moving to multi-feature scenarios to facilitate system validation, considering the complexities of multi-feature datasets and the problem of multiple test correction.

\subsubsection{Varying Sample Sizes}
We began with single-feature scenarios and by experimenting with different tuple-based half-lives $n_{1/2}$. The half-life $n_{1/2}$ parameter in turn controls the size of the samples we make in the batch phase and the streaming EMA tail size. We set a half-life $n_{1/2}$ value and then multiply it by a constant factor of 4 to get the size of the samples to make. We process $4n_{1/2}$ events, meaning any further data we processed would make our oldest events have less than 6\% weight, following Algorithm \ref{pseudo:ema-hist}'s successively applied suppression factor. Processing $4n_{1/2}$ events means the oldest event's contribution is $({2^{- \frac{1}{n_{1/2}}}})^{4n_{1/2}} = {2^{- \frac{4n_{1/2}}{n_{1/2}}}} = 2^{-4} \approx 0.06 = 6\%$.

All the synthetic datasets used throughout this section are a time-series where at each time-step we sample uniformly from a fixed distribution, \textit{i.e.}, there are no time correlations in the used datasets. The reference dataset R1 used in this first experiment contained a single feature that followed a normal distribution with mean $\mu=10$ and standard deviation $\sigma=2$ for the 5 million rows of the dataset. Figure \ref{fig:timeseries-r1} shows the time-series of the single feature in dataset R1. The dataset processed in the streaming phase (\textit{i.e.}, the target dataset) was generated with a size of 5 million events and a single feature. For the first half, \textit{i.e.}, for the first 2.5 million events, the feature followed a normal distribution with mean $\mu=10$ and standard deviation $\sigma=2$, like the reference dataset. For the other half, we changed the generating distribution to be a continuous uniform one with lower bound $a=100$ and upper bound $b=200$. Figure \ref{fig:timeseries-t1} shows the time-series of the single feature in the target dataset T1 where we see the abrupt change in feature values.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/01-reference.pdf}
      \caption{Time-series for dataset R1}
      \label{fig:timeseries-r1}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/01-target.pdf}
      \caption{Time-series for dataset T1}
      \label{fig:timeseries-t1}
    \end{center}
\end{figure}
We conducted three different experiments varying the tuple-based half-life $n_{1/2}$, using the reference dataset R1 and the target dataset T1. The experiment identifiers, the corresponding half-lives and the execution times are shown in Table \ref{tbl:tests-half-life}. Notice that $Sample Size$ corresponds to the size of each sample made in the batch phase and so it is four times the half-life. We used 100 bins (plus the two special infinity bins) for the reference histogram and made 1000 samples.
\begin{table}[!htb]
    \begin{center}
        \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multicolumn{1}{|l|}{\textbf{Experiment}} & \multicolumn{1}{l|}{\textbf{Half-life}} & \multicolumn{1}{l|}{\textbf{Sample Size}} & \multicolumn{1}{l|}{\textbf{Batch execution time}} \\ \hline
        01  & 625  & 2500 & 152                                                          \\ \hline
        02                                        & 62,500                                  & 250,000                               & 2,822                                                        \\ \hline
        03                                        & 250,000                                 & 1,000,000                           & 10,082                                                       \\ \hline
        \end{tabular}
    \end{center}
    \caption{Experiments with varying half-lives}
    \label{tbl:tests-half-life}
\end{table}
Table \ref{tbl:tests-half-life} shows the execution time for the batch phase in seconds. All experiments used 300 Spark executor instances. Since our sample size depends on the half-life (four times larger), the larger the half-life, the larger the sample size. Larger samples imply longer processing times when computing the EMA-like histogram for each sample. With no surprise, we see that the larger the sample the longer it takes to finish the batch phase.

From a functional point of view, were we able to detect the abrupt change introduced in the target dataset T1? Recall that from the 2.5 millionth event forward we change the distribution from a gaussian to an uniform one (Figure \ref{fig:timeseries-t1}). Figure \ref{fig:JSD-signal-01} shows the distance (JSD) values computed during streaming at each event of the 5 million events from dataset T1, using the outputs from the batch phase of experiment 01. We clearly see a change in JSD values, and we move from values close to 0 to values close to 1 (or even 1). Figure \ref{fig:JSD-signal-zoom-01} shows the same signal but for the first half of the streaming period (essentially a zoom-in). We see our JSD values before the change were very low (magnitudes of $10^{-2}$).
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-625.pdf}
      \caption{Whole time-series JSD signal and threshold}
      \label{fig:JSD-signal-01}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-zoom-625.pdf}
      \caption{Zoom-in on JSD signal and threshold}
     \label{fig:JSD-signal-zoom-01}
    \end{center}
\end{figure}
However, the change in JSD value is very abrupt as well. Almost instantly we go from close to 0 values to 1. We also had some false-positives, \textit{i.e.}, some JSD values above the threshold. This indicates to us that the window used is too small and sensitive to small changes, or that we can choose a more strict threshold.

In experiment 02, we use larger sampling windows. Remember our target sliding window in streaming is of the same size as the sampling windows. In the second experiment, we increase the size of the sampling windows by 100 times. Figure \ref{fig:JSD-signal-02} shows the JSD signal during streaming for dataset T1, using the outputs from the batch phase of experiment 02. Again, we see a change in signal from the middle of the streaming period forward. Figure \ref{fig:JSD-signal-zoom-02} shows the same signal for the first half of the streaming period.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-62500.pdf}
      \caption{Whole time-series JSD signal and threshold}
      \label{fig:JSD-signal-02}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-zoom-62500.pdf}
      \caption{Zoom-in on JSD signal and threshold}
     \label{fig:JSD-signal-zoom-02}
    \end{center}
\end{figure}
Notice that this time the growth of our signal is slower. Notice also that we do not have false positives in the first half of the signal. This is because we use a larger window, which means each event has a smaller contribution to the aggregation state. Hence, outliers will have less impact on the aggregation and corresponding signal value.

In experiment 03, we increase the sampling and target windows size once again, this time four times larger than in experiment 02. We would expect an even smoother growth of our signal and, again, no false positives. And indeed, judging by the corresponding signal computation and zoom-in, Figures \ref{fig:JSD-signal-03} and \ref{fig:JSD-signal-zoom-03}, respectively, the signal looks smoother and we see no false positives.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-250000.pdf}
      \caption{Whole time-series JSD signal and threshold}
      \label{fig:JSD-signal-03}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-zoom-250000.pdf}
      \caption{Zoom-in on JSD signal and threshold}
     \label{fig:JSD-signal-zoom-03}
    \end{center}
\end{figure}

\subsubsection{Multiple Distribution Changes}
In this experiment, we wanted to see how our signal behaved when there were multiple change points, where each gaussian distribution was a scaled down version of the previous one. Figure \ref{fig:timeseries-r2} shows the time-series for the reference dataset R2 used in this experiment. It contained 5 million events with a single feature following a gaussian distribution with mean $\mu=100$ and standard deviation $\sigma=50$. The time-series for target dataset T2 is shown in Figure \ref{fig:timeseries-t2}. Notice that we change distributions at 2, 3 and 4 millionth events. Notice also that each subsequent distribution produces values whose domain is a subset of the previous distribution.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/timeseries-r2.pdf}
      \caption{Time-series for dataset R2}
      \label{fig:timeseries-r2}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/timeseries-t2.pdf}
      \caption{Time-series for dataset T2}
     \label{fig:timeseries-t2}
    \end{center}
\end{figure}

The target dataset T2 begins with the same distribution as the reference dataset R2, a gaussian distribution with mean $\mu=100$ and standard deviation $\sigma=50$. At the 2 millionth event mark, we use another Gaussian distribution, this time with mean $\mu=100$ and standard deviation $\sigma=30$. At the 3 millionth event, we change the standard deviation to be $\sigma=10$. From the 4 millionth event to the end, we use a standard deviation of $\sigma=5$.

In this experiment, we used a half-life $n_{1/2}=62500$ and corresponding sample size of 250,000 events, like in the previous section's experiment 02. Figure \ref{fig:JSD-signal-test02} shows the computed JSD signal for the target dataset and Figure \ref{fig:JSD-signal-zoom-test02} zooms in on the first 2 millionth events, before any distribution is changed. We observe that once again we have no false positives in the first 2 million events, which is ideal. 
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-test02.pdf}
      \caption{Whole time-series JSD signal and threshold}
      \label{fig:JSD-signal-test02}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-zoom-test02.pdf}
      \caption{Zoom-in on first 2 million events}
     \label{fig:JSD-signal-zoom-test02}
    \end{center}
\end{figure}
Furthermore, we observe that the computed JSD signal varies accordingly with the changes in distribution. Despite us changing the distributions to subsets of the previous ones to try and deceive our method, our signal accurately represents these changes. In Figure \ref{fig:JSD-signal-test02}, we see several growth points that correspond to the points where the generating gaussian distribution parameters changed. Hence, we are able to detect changes in distribution parameter changes and quantify how big the change is (the closer to 1 our signal is).

\subsubsection{Multiple Distribution Changes, each a Superset of the Previous One}
In this experiment, we wanted to do the reverse of the previous one. In this one, there are multiple change points but each new distribution spans a domain that is a superset of the previous one. Figure \ref{fig:timeseries-r3} shows the reference period used, with 5 million events and one feature following a gaussian distribution of mean $\mu=100$ and standard deviation $\sigma=10$. Figure \ref{fig:timeseries-t3} shows the target period used, also with 5 million events and one feature. Like in the previous experiment, we change distributions at the 2, 3 and 4 millionth event mark. On the other hand, unlike the previous experiment, we now make subsequent distributions cover the previous ones. We change the reference gaussian used with mean $\mu=100$ and standard deviation $\sigma=10$ at the 2, 3 and 4 millionth mark, to use standard deviations $\sigma=30$, $\sigma=40$ and $\sigma=50$, respectively.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/timeseries-r3.pdf}
      \caption{Time-series for dataset R3}
      \label{fig:timeseries-r3}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/timeseries-t3.pdf}
      \caption{Time-series for dataset T3}
     \label{fig:timeseries-t3}
    \end{center}
\end{figure}

Figure \ref{fig:JSD-signal-test03} shows the JSD signal for the target dataset in this experiment. Once again, we were able to detect the changes introduced at the 2, 3 and 4 millionth events. We are also able to identify bigger distribution changes, relative to the reference, by observing bigger JSD values. Figure \ref{fig:JSD-signal-zoom-test03} shows a zoomed in portion of the original signal, where we can verify that there are no false positives.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-test03.pdf}
      \caption{Whole time-series JSD signal and threshold}
      \label{fig:JSD-signal-test03}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-zoom-test03.pdf}
      \caption{Zoom-in on first 2 million events}
     \label{fig:JSD-signal-zoom-test03}
    \end{center}
\end{figure}

\subsubsection{Return to Normality After Change}
Our goal with this experiment was to check if our signal returned to its normal values (below the alert threshold) if we changed our distribution to the original reference one, as intended.

In this experiment, we used reference dataset R4 (Figure \ref{fig:timeseries-r4}) and target dataset T4 (Figure \ref{fig:timeseries-t4}). 
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/timeseries-r4.pdf}
      \caption{Time-series for dataset R4}
      \label{fig:timeseries-r4}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/timeseries-t4.pdf}
      \caption{Time-series for dataset T4}
     \label{fig:timeseries-t4}
    \end{center}
\end{figure}

The reference dataset R4 had a single feature that followed a Gaussian distribution of $\mu=100$ and $\sigma=5$ for all 5 million events. Up until the 2 millionth event, the target dataset T4 followed a Gaussian distribution of $\mu=100$ and $\sigma=5$, similarly to the reference dataset R4. At the 2 millionth event, we change the Gaussian parameters to $\mu=100$ and $\sigma=30$. At the 3 millionth mark, we once again change the Gaussian parameters to $\mu=100$ and $\sigma=50$. Finally, on the 4 millionth event, we return to our reference distribution, a gaussian with $\mu=100$ and $\sigma=5$.

Figure \ref{fig:JSD-signal-test04} shows the computed JSD signal for the target period. As expected, our JSD signal increases at the 2 millionth and 3 millionth event mark, because the distribution parameters change. At the 4 millionth event, we use the original reference distribution and notice that the JSD signal returns to values below the threshold, considering it a normal state, as we intended. Figure \ref{fig:JSD-signal-zoom-test04} is a zoom in on the first 2 millionth events before any change occurs, and we see no false positives.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-test04.pdf}
      \caption{Whole time-series JSD signal and threshold}
      \label{fig:JSD-signal-test04}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/stream-analysis-viz-zoom-test04.pdf}
      \caption{Zoom-in on first 2 million events}
    \label{fig:JSD-signal-zoom-test04}
    \end{center}
\end{figure}

\subsubsection{An Attempt to Deceive our Method}
In this experiment, we attempt to deceive our method by using a different distribution from the reference period but ensuring its domain is roughly the same as the reference one.

We used the reference dataset R5 (Figure \ref{fig:timeseries-r5}) which had 5 million events, with a single feature that followed a gaussian with  $\mu=100$ and $\sigma=10$. Our target dataset T5 is represented in Figure \ref{fig:timeseries-t5}. For the first 2 million events, the generating distribution used was similar to the one used in the reference dataset, a gaussian with $\mu=100$ and $\sigma=10$. However, at the 2 millionth event, we change the gaussian standard deviation to $\sigma=30$. Then, at the 4 millionth event, we change the distribution type. We use an uniform distribution with lower bound $a=80$ and upper bound $b=120$. Note in Figure \ref{fig:timeseries-t5} that this last portion resembles the first one, the reference.
\begin{center}
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/timeseries-r5.pdf}
  \captionof{figure}{Time-series for dataset R5}
  \label{fig:timeseries-r5}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/timeseries-t5.pdf}
  \captionof{figure}{Time-series for dataset T5}
  \label{fig:timeseries-t5}
\end{minipage}
\end{center}

Let's take a look at the signal computed for the target period. Figure \ref{fig:JSD-signal-test05} shows us the JSD signal for dataset T5 (Figure \ref{fig:timeseries-t5}). As expected, we see a rapid growth of JSD value after the 2 millionth event, when we change the Gaussian parameters. We then change the distribution type, from a gaussian to a uniform one. Despite the latter having a domain contained within the former, we see that our signal is accurate and does not drop below the alert threshold. This is still a different distribution that we want to alert. Figure \ref{fig:JSD-signal-zoom-test05} is a zoom in of the first 2 millionth events and once again we have no false positives.
\begin{figure}[!htb]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/stream-analysis-viz-test05.pdf}
  \captionof{figure}{JSD signal and threshold}
  \label{fig:JSD-signal-test05}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{figures/stream-analysis-viz-zoom-test05.pdf}
  \caption{Zoom in on the first 2 million events}
  \label{fig:JSD-signal-zoom-test05}
\end{subfigure}
\caption{JSD signal and threshold}
\end{figure}

\subsection{Multi-Feature Analysis}
Before moving to real data, we performed one last experiment, using synthetic data but this time performing a multi-feature analysis. Table \ref{tbl:multi-feat-ref-dataset-distros} shows the underlying generating distributions and parameters for each of the four features ($x_1$, $x_2$, $x_3$ and $x_4$) in the reference multi-feature dataset R6. Figures \ref{fig:timeseries-r6-x1}, \ref{fig:timeseries-r6-x2}, \ref{fig:timeseries-r6-x3} and \ref{fig:timeseries-r6-x4} show the reference time-series for features $x_1$, $x_2$, $x_3$ and $x_4$ on the reference dataset R6, respectively.

\begin{table}[!htb]
    \begin{center}
        \begin{tabular}{|c|c|c|ll}
        \cline{1-3}
        \textbf{Feature} & \textbf{Reference Distribution} & \textbf{Parameters} &  &  \\ \cline{1-3}
        $x_1$               & Gaussian/Normal                 & $\mu=100$, $\sigma=5$  &  &  \\ \cline{1-3}
        $x_2$               & Gaussian/Normal                 & $\mu=100$, $\sigma=5$  &  &  \\ \cline{1-3}
        $x_3$               & Uniform                         & $a=200$, $b=250$       &  &  \\ \cline{1-3}
        $x_4$               & Uniform                         & $a=50$, $b=125$        &  &  \\ \cline{1-3}
        \end{tabular}
    \end{center}
    \caption{Multi-feature reference dataset and feature distributions}
    \label{tbl:multi-feat-ref-dataset-distros}
\end{table}
